Postural developments modulate children's visual access to social information

### 2018-05-11 writeup-cogsci/cogsci2018_cameraReady.Rmd
cogsci revisions

Then, however, their view of the world radically changes [@adolph2006motor]. Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently. These motor changes have significant consequences for what children see; crawling and walking infants simply have different views of the world. For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]. In general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].

These motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities in various ways. For example, postural changes impact how children interact with their mothers: walking (vs. crawling) infants make different kinds of object-related bids for attention from their mothers and tend to hear more action directed statements (e.g., “open it”) [@karasik2014]. Infants also increase interactive behaviors in general and explore their surroundings more [@thurman2017spatial]. Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies. Thus, children’s ability to independently locomote may change their ability to access social information (e.g., faces, gaze) and in turn accelerate their learning.

@@

While it is uncertain how much these changes are driven by the concurrent motoric changes described above versus by other developmental factors, recent work has found that whether an infant is crawling, sitting, or standing predicts how often they gaze at faces and bodies [@franchak2017see].

Here, we directly examine whether postural and locomotor developments change the availability of social information -- the presence of faces and hands. To do so, we recorded the visual experience of a group of infants in three age ranges (8, 12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated. We then capitalized on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers. Only one piece of work to date has directly examined the impact of postural development to social information access in infants: a study by @franchak2017see found support for the motoric hypothesis presented above using a similar free-play experimental paradigm with head-mounted cameras on infants. This study required manual coding of the first 5 minutes of 15 minute videos from a cohort of 12-month olds; here we replicate and extend their findings using novel computer vision techniques to allow for video analysis at scale. We also analyze infants from a larger developmental age range (8, 12, and 16 month-olds).

We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus providing children with greater access to the social information in their environment.

@@

# initial submission data - in case you want to reproduce.
# d <- read_feather("../data/consolidated_data_4detectors.feather") %>%
#   mutate(posture = factor(posture), orientation = factor(orientation))

@@

# Function to evaluate detectors
evaluate <- function(a, b) {
  if (a == TRUE) {
    if (a == b) return ("TP") # was face/wrist, detected face/wrist
    else return("FN") # was face/wrist, missed face/wrist
  }
  else {
    if (a == b) return("TN") # was not face/wrist, detected face/wrist
    else return("FP") # was not face/wrist, said face/wrist
  }
}
```

```{r  include=FALSE, echo=FALSE}
## Evalute face and wrist detectors
performanceFaces <- ground_truth_faces  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong)   %>%
  filter(detectorType!="OpenPose-Wrists") %>%
  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face))

@@

## EVALUATE WRIST DETECTOR
performanceWrists <- ground_truth_wrists  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong) %>%
  filter(detectorType=="OpenPose-Wrists") %>%
  mutate(detector_is_wrist = detection, is_wrist = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_wrist, detector_is_wrist))

@@

#wrangle data
dGLM_faces <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1]) %>%
  mutate(propFaces = countFaces / (countFaces + countNotFaces))

dGLM_wrists <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE),     age.at.test=age.at.test[1])  %>%
  mutate(propWrists = countWrists / (countWrists + countNotWrists))

## run models
faceGLM <- glmer(cbind(countFaces,countNotFaces) ~ orientation + posture + scale(age.at.test) + 
               (orientation | subid), 
             dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

@@

## output
cleanedNames_NoInt = c("Intercept","Close","Far","Sit","Stand","Age") # march 6th - prone is dropped from GLMs
# cleanedNames_NoInt = c("Intercept","Close","Far","Prone","Sit","Stand","Age") # from initial submission

@@

```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
detectionsByAge <- dlong %>%
  filter(detectorType!="OpenPose-Faces") %>% # for brevity
  filter(detectorType!="ViolaJones-Faces") %>% # low performance
  filter(!is.na(age.at.test)) %>%
  group_by(subid, age.at.test, detectorType) %>%
  summarise(detections = mean(detection)) %>%
  left_join(inc_subs) %>%
  mutate(len = as.numeric(len.min))

@@

aroundNamings_Posture_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, posture) %>%
  summarize(propFacesDetected = mean(detections)) %>%
  filter(!is.na(propFacesDetected)) %>%
  filter(!is.na(posture)) %>%
  group_by(posture) %>%
  multi_boot_standard(col = "propFacesDetected")

@@

## Wrists
wrists_baseline<- d %>%
  group_by(subid, age.grp) %>%
  summarize(baselineDetections = mean(faceMT))
#
wrists_naming <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, age.grp)  %>%
  summarize(namingDetections = mean(detections, na.rm=TRUE)) %>%
  left_join(wrists_baseline) %>%
  mutate(baselineDiff = namingDetections - baselineDetections)  %>%
  group_by(age.grp) %>%
  summarize(avgBaselineDiff = mean(baselineDiff))

>>>>

Then, however, their view of the world radically changes [@adolph2006motor]. Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently. These motor changes have significant consequences for what children see. For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].

These motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities in various ways. For example, postural changes impact how children interact with their mothers: walking (vs. crawling) infants make different kinds of object-related bids for attention from their mothers, tend to hear more action directed statements (e.g., “open it”), and tend to explore their surroundings more [@karasik2014; @thurman2017spatial]. In an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies. Thus, children’s ability to independently locomote may change their ability to access social information (e.g., faces, gaze) and in turn accelerate their learning.

@@

Here, we directly examine whether postural and locomotor developments change the availability of social information -- the presence of faces and hands. We recorded the visual experience of a group of infants in three age ranges (8, 12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated. Recent work directly examined how postural developments change access to social information in 12 month-olds in a similar paradigm, finding support for the motoric hypothesis [@franchak2017see]. Here we replicate and extend their main finding using novel computer vision techniques to allow for video analysis at scale. In particular, we capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers. We predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus providing children with greater access to the social information in their environment.

@@

@@

# Function to evaluate detectors
evaluate <- function(a, b) {
  if (a == TRUE) {
    if (a == b) return ("TP") # was face/wrist, detected face/wrist
    else return("FN") # was face/wrist, missed face/wrist
  }
  else {
    if (a == b) return("TN") # was not face/wrist, did not detect face/wrist
    else return("FP") # was not face/wrist, detected face/wrist
  }
}
```

```{r  include=FALSE, echo=FALSE}
## Evalute face and wrist detectors
performanceFaces <- ground_truth_faces  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong)   %>%
  filter(detectorType!="OpenPose-Wrists") %>%
  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face)) # always have the gronud truth first in evaluate argument

@@

## EVALUATE WRIST DETECTOR
performanceWrists <- ground_truth_wrists  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong) %>%
  filter(detectorType=="OpenPose-Wrists") %>%
  mutate(detector_is_wrist = detection, is_wrist = as.logical(is_wrist)) %>% 
  rowwise() %>%
  mutate(result = evaluate(is_wrist, detector_is_wrist))

@@

#wrangle data
## Using MTCNN detections for GLM since it was the best
dGLM_faces <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

dGLM_wrists <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

## run models 
faceGLM <- glmer(cbind(countFaces,countNotFaces) ~ posture + orientation + scale(age.at.test) + 
               (orientation | subid), 
             dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

@@

## output
cleanedNames_NoInt = c("Intercept","Close","Far","Sit","Stand","Age")

@@

```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
detectionsByAge <- dlong %>%
  filter(detectorType!="OpenPose-Faces") %>% # for brevity
  filter(detectorType!="ViolaJones-Faces") %>% # low performance
  filter(!is.na(age.at.test)) %>%
  group_by(subid, age.at.test, detectorType) %>%
  summarise(detections = mean(detection)) %>%
  left_join(inc_subs) %>%
  mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video

@@

aroundNamings_Posture_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceMT) %>% ## for covergence with helper.R file, which looks for this variable name
  do(summarize.naming(.)) %>%
  group_by(subid, posture) %>%
  summarize(propFacesDetected = mean(detections)) %>%
  filter(!is.na(propFacesDetected)) %>%
  filter(!is.na(posture)) %>%
  group_by(posture) %>%
  multi_boot_standard(col = "propFacesDetected")

@@

## Wrists
wrists_baseline<- d %>%
  group_by(subid, age.grp) %>%
  summarize(baselineDetections = mean(wristOP))
#
wrists_naming <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.)) %>%
  group_by(subid, age.grp)  %>%
  summarize(namingDetections = mean(detections, na.rm=TRUE)) %>%
  left_join(wrists_baseline) %>%
  mutate(baselineDiff = namingDetections - baselineDetections)  %>%
  group_by(age.grp) %>%
  summarize(avgBaselineDiff = mean(baselineDiff))

### 2018-05-11 writeup-cogsci/cogsci2018_cameraReady.tex
cogsci revisions

Then, however, their view of the world radically changes (Adolph \&
Berger, 2006). Infants' motor abilities improve dramatically near the
end of the first year of life, allowing them to locomote independently.
These motor changes have significant consequences for what children see;
crawling and walking infants simply have different views of the world.
For example, during spontaneous play in a laboratory playroom, toddlers
are more likely to look at the floor while crawling than while walking
(Franchak, Kretch, Soska, \& Adolph, 2011). In general, walking infants
tend to have full visual access to their environment and the people in
it, while crawling infants do not (Kretch, Franchak, \& Adolph, 2014).

These motor improvements lead to developmental cascades that impact
children's emerging social, cognitive, and linguistic abilities in
various ways. For example, postural changes impact how children interact
with their mothers: walking (vs.~crawling) infants make different kinds
of object-related bids for attention from their mothers and tend to hear
more action directed statements (e.g., ``open it'') (Karasik,
Tamis-LeMonda, \& Adolph, 2014). Further, in an observational study,
Walle \& Campos (2014) found that children who were able to walk had
both higher receptive and productive vocabularies. Thus, children's
ability to independently locomote may change their ability to access
social information (e.g., faces, gaze) and in turn accelerate their
learning.

Recent technological developments allow for testing of this hypothesis
by documenting the experiences of infants and children from their own
perspective. By using head-mounted cameras, researchers have begun to
record the visual experiences of infants and children -- which even for
walking children are extremely different from the adult perspective (and
not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
Yu, \& Smith, 2017; Franchak et al., 2011; Yoshida \& Smith, 2008).
Children's views tend to be more restricted and dominated by objects and
hands (Yoshida \& Smith, 2008), and computational and empirical work
suggests that this restricted viewpoint may be more effective for
learning objects and their labels than the comparable adult perspective
(Bambach, Crandall, Smith, \& Yu, 2017; Yurovsky, Smith, \& Yu, 2013).
This perspective changes over the first two years of life, as visual
input transitions from primarily containing close up views of faces to
capturing views of hands paired with the objects they are acting on
(Fausey, Jayaraman, \& Smith, 2016). It is uncertain how much these
changes are driven by the concurrent motoric changes described above
versus by other developmental factors.

Here, we directly examine whether postural and locomotor developments
change the availability of social information -- the presence of faces
and hands. To do so, we recorded the visual experience of a group of
infants in three age ranges (8, 12, and 16 months) using head-mounted
cameras during a brief laboratory free-play session; children's posture
and orientation relative to their caregiver were also recorded from a
third-person perspective and hand-annotated. We then capitalize on
recent improvements in face and pose detection algorithms (Cao, Simon,
Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
frequencies of faces and hands (using wrists as a proxy for the latter)
in the child's visual environment, both overall and relative to naming
events by their caregivers. We hypothesized that there would be
differential access to social information based on children's postural
developments: crawling infants would see fewer faces because they would
primarily be looking at the ground, while walking toddlers would have
access to a richer visual landscape, thus providing children with
greater access to the social information in their environment.

@@

Headcam videos were trimmed such that they excluded the instruction
phase when the experimenter was in the room and were automatically
synchronized with the tripod-mounted videos using FinalCut Pro Software.
These sessions yielded 516 minutes (almost a million frames) of video,
with an average video length of 14.33 minutes (min = 4.53, max = 19.35).

@@

Finally, we explored how face and hand detections changed during object
labeling events as a function of infants' posture and orientation. We
analyzed a four-second window around each labeling event (e.g., ``Look
at the {[}zem{]}!''); these labeling events were hand-annotated and
synchronized with the frame-by-frame face/hand detections. We found that
infants' posture and orientations modulated their visual access to faces
and hands during labeling events; infants who were sitting or standing
were more likely to have visual access to this social information (see
Figure \ref{fig:detByNaming}). However, we did not find that infants saw
particularly more faces or hands during naming events relative to
baseline (avg. difference in proportion of hands, 8 m.o. = 0, 12 m.o. =
-0.004, 16 m.o. = -0.014; avg. difference in proportion of faces, 8 m.o.
= 0.003, 12 m.o. = 0.01, 16 m.o. = 0.018).

@@

We used a head-mounted camera to explore how children's postural and
locomotor development affects their visual access to social information,
here operationalized as the presence of the faces and hands of their
caregiver. Children's posture and orientation towards their caregiver
changed systematically across age, and both of these factors influenced
the proportion of faces and hands that were available in the child's
visual field. This work suggests that motor development modulates how
infants experience their visual world and the social information in it:
Infants that are sitting and standing have a different view of their
world, the people in it, and the actions that are being performed.

@@

@@

>>>>

Then, however, their view of the world radically changes (Adolph \&
Berger, 2006). Infants' motor abilities improve dramatically near the
end of the first year of life, allowing them to locomote independently.
These motor changes have significant consequences for what children see.
For example, during spontaneous play in a laboratory playroom, toddlers
are more likely to look at the floor while crawling than while walking
(Franchak, Kretch, Soska, \& Adolph, 2011); in general, walking infants
tend to have full visual access to their environment and the people in
it, while crawling infants do not (Kretch, Franchak, \& Adolph, 2014).

These motor improvements lead to developmental cascades that impact
children's emerging social, cognitive, and linguistic abilities in
various ways. For example, postural changes impact how children interact
with their mothers: walking (vs.~crawling) infants make different kinds
of object-related bids for attention from their mothers, tend to hear
more action directed statements (e.g., ``open it''), and tend to explore
their surroundings more (Karasik, Tamis-LeMonda, \& Adolph, 2014;
Thurman \& Corbetta, 2017). In an observational study, Walle \& Campos
(2014) found that children who were able to walk had both higher
receptive and productive vocabularies. Thus, children's ability to
independently locomote may change their ability to access social
information (e.g., faces, gaze) and in turn accelerate their learning.

Recent technological developments allow for testing of this hypothesis
by documenting the experiences of infants and children from their own
perspective. By using head-mounted cameras, researchers have begun to
record the visual experiences of infants and children -- which even for
walking children are extremely different from the adult perspective (and
not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
Yu, \& Smith, 2017; Franchak et al., 2011; Yoshida \& Smith, 2008).
Children's views tend to be more restricted and dominated by objects and
hands (Yoshida \& Smith, 2008), and computational and empirical work
suggests that this restricted viewpoint may be more effective for
learning objects and their labels than the comparable adult perspective
(Bambach, Crandall, Smith, \& Yu, 2017; Yurovsky, Smith, \& Yu, 2013).
This perspective changes over the first two years of life, as visual
input transitions from primarily containing close up views of faces to
capturing views of hands paired with the objects they are acting on
(Fausey, Jayaraman, \& Smith, 2016).

Here, we directly examine whether postural and locomotor developments
change the availability of social information -- the presence of faces
and hands. We recorded the visual experience of a group of infants in
three age ranges (8, 12, and 16 months) using head-mounted cameras
during a brief laboratory free-play session; children's posture and
orientation relative to their caregiver were also recorded from a
third-person perspective and hand-annotated. Recent work directly
examined how postural developments change access to social information
in 12 month-olds in a similar paradigm, finding support for the motoric
hypothesis (Franchak, Kretch, \& Adolph, 2017). Here we replicate and
extend their main finding using novel computer vision techniques to
allow for video analysis at scale. In particular, we capitalize on
recent improvements in face and pose detection algorithms (Cao, Simon,
Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
frequencies of faces and hands (using wrists as a proxy for the latter)
in the child's visual environment, both overall and relative to naming
events by their caregivers. We predicted that there would be
differential access to social information based on children's postural
developments: crawling infants would see fewer faces because they would
primarily be looking at the ground, while walking toddlers would have
access to a richer visual landscape, thus providing children with
greater access to the social information in their environment.

@@

Headcam videos were trimmed such that they excluded the instruction
phase when the experimenter was in the room and were automatically
synchronized with the tripod-mounted videos using FinalCut Pro Software.
These sessions yielded 507 minutes (almost a million frames) of video,
with an average video length of 14.07 minutes (min = 4.53, max = 19.35).

@@

Finally, we explored how face and hand detections changed during object
labeling events as a function of infants' posture and orientation. We
analyzed a four-second window around each labeling event (e.g., ``Look
at the {[}zem{]}!''); these labeling events were hand-annotated and
synchronized with the frame-by-frame face/hand detections. We found that
infants' posture and orientations modulated their visual access to faces
and hands during labeling events; infants who were sitting or standing
were more likely to have visual access to this social information (see
Figure \ref{fig:detByNaming}). However, we did not find that infants saw
particularly more faces or hands during naming events relative to
baseline (avg. difference in proportion of hands, 8 m.o. = 0.012, 12
m.o. = 0.012, 16 m.o. = 0.013; avg. difference in proportion of faces, 8
m.o. = 0.003, 12 m.o. = 0.01, 16 m.o. = 0.018).

@@

We used a head-mounted camera to explore how children's postural and
locomotor development affects their visual access to social information,
here operationalized as the presence of the faces and hands of their
caregiver. Children's posture and orientation towards their caregiver
changed systematically across age, and both of these factors influenced
the proportion of faces and hands that were available in the child's
visual field. Though we oriented our cameras to maximally capture face
information, we nonetheless saw an influence of orientation and posture
on visual access to hands. This work suggests that motor development
modulates how infants experience their visual world and the social
information in it: Infants that are sitting and standing have a
different view of their world, the people in it, and the actions that
are being performed.

@@

\hypertarget{ref-franchak2017see}{}
Franchak, J. M., Kretch, K. S., \& Adolph, K. E. (2017). See and be
seen: Infant--caregiver social looking during locomotor free play.
\emph{Developmental Science}.

@@

\hypertarget{ref-thurman2017spatial}{}
Thurman, S. L., \& Corbetta, D. (2017). Spatial exploration and changes
in infant--mother dyads around transitions in infant locomotion.
\emph{Developmental Psychology}, \emph{53}(7), 1207.
